---
title: "Homework04:Tidy data and joins"
output: 
  html_document:
    keep_md: TRUE
  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### Load packages
```{r load_pkg}
suppressPackageStartupMessages(library(tidyverse))
library(knitr)
library(gapminder)
library(kableExtra)
```

### Data Reshaping
For data reshaping, I am using a portion of gapminder data.
I will do a mix of Activity #1 and #2 on this data

#### Data
```{r data_dr}
data_dr <- gapminder %>% 
            filter(country %in% c("Singapore", 
                                  "Italy",
                                  "Russia",
                                  "India"),
                   year > 1951 & year < 1972) %>%
            select(continent,country,year,lifeExp)
kable(data_dr) %>% 
  kable_styling(c("striped", "bordered"),
                full_width = F)
```

Converting the data for each country to show the yearly trend in one row(columns for each year)

```{r spread_dr1}
sdr <- data_dr %>%
        spread(year, lifeExp)
kable(sdr) %>% 
  kable_styling(c("striped", "bordered"),
                full_width = F)
```

The above table is easy to analyse and mutate. For example if want to add a column for change in lifeExp from 1952 to 1967:

```{r mutate_dr}
sdr %>%
  mutate(change = `1967` - `1952`) %>%
  kable() %>% 
  kable_styling(c("striped", "bordered"),
                full_width = F)
```

To revert the changes:

```{r gather_sdr}
sdr %>% 
  gather(year,lifeExp,"1952","1957","1962","1967") %>%
  kable() %>% 
  kable_styling(c("striped", "bordered"),
                full_width = F)
```

Now lets spread the initial data to have separate columns for each country

```{r spread_dr2}
data_dr %>% 
  spread(country,lifeExp) %>% 
  kable() %>% 
  kable_styling(c("striped", "bordered"),
                full_width = F)
```

Lets make a scatterplot for lifeExp of India vs Singapore

#### Data of two countries:

```{r scatter_data}
data_dr %>%
  filter(continent == "Asia") %>%
  spread(country,lifeExp) %>%
  kable() %>% 
  kable_styling(c("striped", "bordered"),
                full_width = F)
```


#### Scatter-plot

```{r scatter}
data_dr %>%
  filter(continent == "Asia") %>%
  spread(country,lifeExp) %>%
  ggplot(aes(India,Singapore)) +
  geom_point(aes(color = year,
                 cex=.02)) +
  ggtitle("LifeExp India vs Singapore")

```

So, plot above shows that there is almost linear increase in lifeExp for both the countries.

### Table join
For table join cheatsheet I am doing Activity #2

#### Data
I am taking data of some of the shows(I like) and their ratings on IMDb.

```{r data_def}
shows <- tibble(name = c("Office","Walking Dead",
                         "Stranger Things","Friends"), 
                genre = c("comedy","Sci-fi",
                          "Sci-fi","comedy"),
                continuing = c("No","Yes","Yes","No"),
                seasons = c(10, 9, 2, 10))
rating <- tibble(name = c("Walking Dead", "Office",
                          "Narcos", "Stranger Things"), 
                 rating = c(8.4, 8.8, 9.1, 9.0))
kable(shows, caption = "Shows") %>% 
  kable_styling(c("striped", "bordered"),
                full_width = F)
kable(rating, caption = "Rating") %>% 
  kable_styling(c("striped", "bordered"),
                full_width = F)

```

Lets try some of the table joining functions on tables above.

If I want to combine the data from above two tables for my analysis, I can use inner_join(). But if I want to discard the observations from one table which are not there in my second table without mutating the columns, I can use semi_join()

```{r join1}
ijsr <- inner_join(shows, rating)
sjsr <- semi_join(shows,rating)
kable(ijsr, caption = "Inner_join()") %>% 
  kable_styling(c("striped", "bordered"),
                full_width = F)
kable(sjsr, caption = "Semi_join()") %>% 
  kable_styling(c("striped", "bordered"),
                full_width = F)
```

As I did not mention the variable label to be used to join the tables, it does that by the common label found in two tibbles(name). 
Also, as we can see semi_join(x,y) retains the data for x="shows". So, we can define x and y as per our requirement.

In above both the functions data was lost. If we do not want to lose the data from our primary table, we can do so by using either left_join() or right_join()

```{r join2}
ljsr <- left_join(shows, rating)
rjsr <- right_join(shows,rating)
kable(ljsr, caption = "Left_join()") %>% 
  kable_styling(c("striped", "bordered"),
                full_width = F)
kable(shows, caption = "Right_join()") %>% 
  kable_styling(c("striped", "bordered"),
                full_width = F)
```


As you can see, left_join(x,y) retains all rows of x and right_join(x,y) retains all rows of y.

If we have a huge database and want to see which observations from primary table are missing in secondary table, we can use anti_join().

```{r join3}
ajsr1 <- anti_join(shows, rating)
ajsr2 <- anti_join(rating,shows)
kable(ajsr1, caption = "Anti_join(Shows,Rating)") %>% 
  kable_styling(c("striped", "bordered"),
                full_width = F)
kable(ajsr2, caption = "Anti_join(Rating,Shows)") %>% 
  kable_styling(c("striped", "bordered"),
                full_width = F)
```

If we decide that we cannot afford to loose any data, full_join() can be used.

```{r join4}
fjsr <- full_join(shows, rating)
kable(fjsr, caption = "Full_join()") %>% 
  kable_styling(c("striped", "bordered"),
                full_width = F)
```

Lets see what merge() does to the data above.

```{r merge_data}
msr <- merge(shows, rating)
kable(ijsr, caption = "Inner_join()") %>% 
  kable_styling(c("striped", "bordered"),
                full_width = F)
kable(msr, caption = "Merge()") %>% 
  kable_styling(c("striped", "bordered"),
                full_width = F)
```

So, the output of merge() is same as dplyr function inner_join(). However, in inner_join(x,y), the order of output is same as x="shows" but that is not the case in merge().
